---
title: "assignment 3"
author: "Jacob O"
date: "2023-10-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(mosaicData)
```

## Question 1
1. For the following regular expression, explain in words what it matches on. Then add test strings to demonstrate that it in fact does match on the pattern you claim it does. Make sure that your test set of strings has several examples that match as well as several that do not. *If you copy the Rmarkdown code for these exercises directly from my source pages, make sure to remove the `eval=FALSE` from the R-chunk headers.*
    a) This regular expression matches:  a
```{r}
strings <- c('a', 'banana', 'orangutan', 'AAAAA', 'I want a mule', 'no, not the animal', 'you know what i mean', 'from moscow', 'bee', 'more stuff', 'blAh')
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, 'a') )
```
    b) This regular expression matches:  'ab'
```{r}
strings <- c('b', 'a', 'christened', 'bayybeee', 'altruism', 'biscuit', 'now I want a biscoff', 'absent', 'ablative', 'cab')
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, 'ab') )
```
    c)  This regular expression matches:  a or b
```{r}
strings <- c('b', 'a', 'christened', 'bayybeee', 'altruism', 'biscuit', 'now I want a biscoff', 'cortisol', 'length', 'ripper')
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '[ab]') )
```
    d)  This regular expression matches:  a or b at the beginning
```{r}
strings <- c('cab', 'octal', 'obituary', 'apple', 'belch', 'beginning', 'actor')
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '^[ab]') )
```
    e)  This regular expression matches: 1 or more digits, some whitespace, then an a or A
```{r}
strings <- c('6639 ABQ', 'representin tha ABQ', 'my name is skyler white, yo', '1 a', '1a', '123 b', '1     a', '8675309 AAAAAAAAAAAA', 'yeah thats it')
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '\\d+\\s[aA]') )
```
    f)  This regular expression matches:  same as before but the whitespace can be multiple, or 0
```{r}
strings <- c('6639 ABQ', 'representin tha ABQ', 'my name is skyler white, yo', '1 a', '1a', '123 b', '1     a, many spaces', '8675309 AAAAAAAAAAAA', 'yeah thats it')
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '\\d+\\s*[aA]') )
```
    g)  This regular expression matches:  literally anything lmao
```{r}
strings <- c('lighter',
'nut',
'location',
'sculpture',
'dealer',
'lamp',
'wolf',
'copy',
'flawed',
'12348127834')
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '.*') )
```
    h) This regular expression matches: beginning of string, 2 alphanumerics, then 'bar'
```{r}
strings <- c('54bar', 'babar', '9Sbar', 'lalalal 54bar', '32 bar', '6h ba')
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '^\\w{2}bar') )
```
    i) This regular expression matches: 'foo.bar'. or the previous question thing
```{r}
strings <- c('54bar', 'babar', '9Sbar', 'lalalal 54bar', '32 bar', '6h ba', 'foo.bar', 'foobar', 'foo%bar')
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '(foo\\.bar)|(^\\w{2}bar)') )
```

## Question 2
2. The following file names were used in a camera trap study. The S number represents the site, P is the plot within a site, C is the camera number within the plot, the first string of numbers is the YearMonthDay and the second string of numbers is the HourMinuteSecond.
```{r}
file.names <- c( 'S123.P2.C10_20120621_213422.jpg',
                 'S10.P1.C1_20120622_050148.jpg',
                 'S187.P2.C2_20120702_023501.jpg')
```
    Produce a data frame with columns corresponding to the `site`, `plot`, `camera`, `year`, `month`, `day`, `hour`, `minute`, and `second` for these three file names. So we want to produce code that will create the data frame:
    
    ```{r, eval=FALSE}
     Site Plot Camera Year Month Day Hour Minute Second
     S123   P2    C10 2012    06  21   21     34     22
      S10   P1     C1 2012    06  22   05     01     48
     S187   P2     C2 2012    07  02   02     35     01
    ```
    
```{r}
extractorman <- '^([sS]\\d+)\\.([pP]\\d+)\\.([cC]\\d+)_(\\d{4})(\\d{2})(\\d{2})_(\\d{2})(\\d{2})(\\d{2})'
names <- c("Site", "Plot", "Camera", "Year", "Month", "Day", "Hour", "Minute", "Second")
dfman <- NULL
for(i in 1:9) {
  dfman <- cbind(dfman, str_extract(file.names, extractorman, group = i))
}
colnames(dfman) <- names
dfman <- data.frame(dfman)
dfman
```

## Question 3
3. The full text from Lincoln's Gettysburg Address is given below. Calculate the mean word length *Note: consider 'battle-field' as one word with 11 letters*).
```{r}
Gettysburg <- 'Four score and seven years ago our fathers brought forth on this 
continent, a new nation, conceived in Liberty, and dedicated to the proposition 
that all men are created equal.

Now we are engaged in a great civil war, testing whether that nation, or any 
nation so conceived and so dedicated, can long endure. We are met on a great 
battle-field of that war. We have come to dedicate a portion of that field, as 
a final resting place for those who here gave their lives that that nation might 
live. It is altogether fitting and proper that we should do this.

But, in a larger sense, we can not dedicate -- we can not consecrate -- we can 
not hallow -- this ground. The brave men, living and dead, who struggled here, 
have consecrated it, far above our poor power to add or detract. The world will 
little note, nor long remember what we say here, but it can never forget what 
they did here. It is for us the living, rather, to be dedicated here to the 
unfinished work which they who fought here have thus far so nobly advanced. It 
is rather for us to be here dedicated to the great task remaining before us -- 
that from these honored dead we take increased devotion to that cause for which 
they gave the last full measure of devotion -- that we here highly resolve that 
these dead shall not have died in vain -- that this nation, under God, shall 
have a new birth of freedom -- and that government of the people, by the people, 
for the people, shall not perish from the earth.'

yahhhbebe <- str_extract_all(Gettysburg, "(\\w+)-*(\\w+)")
head(yahhhbebe[[1]])
mean(str_length(yahhhbebe[[1]]))
```

## Question 1

1. Convert the following to date or date/time objects.
    a) September 13, 2010.
```{r}
mdy("September 13, 2010")
```
    b) Sept 13, 2010.
```{r}
mdy("Sept 13, 2010")
```
    c) Sep 13, 2010.
```{r}
mdy("Sep 13, 2010")
```
    d) S 13, 2010. Comment on the month abbreviation needs.
```{r}
mdy("S 13, 2010")
```
It seems that you have to do either the full month name, or a 3 letter abbreviation
    e) 07-Dec-1941.
```{r}
dmy("07-Dec-1941")
```
    f) 1-5-1998. Comment on why you might be wrong.
```{r}
mdy("1-5-1998")
```
Day vs month is ambiguous. But that's only for europeans and I'm too busy being a free God-fearing, red-blooded, burger-grilling American. God bless America!
    g) 21-5-1998. Comment on why you know you are correct.
```{r}
dmy("21-5-1998")
```
Day vs month is not ambiguous. there is no month 21.
    h) 2020-May-5 10:30 am
```{r}
ymd_hm("2020-May-5 10:30 am")
```
    i) 2020-May-5 10:30 am PDT (ex Seattle)
```{r}
ymd_hm("2020-May-5 10:30 am", tz="PST8PDT")
```
    j) 2020-May-5 10:30 am AST (ex Puerto Rico)
```{r}
ymd_hm("2020-May-5 10:30 am", tz="America/Puerto_Rico")
```

## Question 2
2. Using just your date of birth (ex Sep 7, 1998) and today's date calculate the following _Write your code in a manner that the code will work on any date after you were born._:
    a) Calculate the date of your 64th birthday.
```{r}
mdy("March 15, 2002") + years(64)
```
    b) Calculate your current age (in years). _Hint: Check your age is calculated correctly if your birthday was yesterday and if it were tomorrow!_ 
```{r}
ageYAAAAAHHHH <- year(as.period(mdy("Mar 15, 2002") %--% ymd(lubridate::today())  ))
ageYAAAAAHHHH
```
^^^^ worlds most readable single line of code ^^^^
    d) Using your result in part (b), calculate the date of your next birthday.
```{r}
birthday<- mdy("Mar 15, 2002")
year(birthday) <- year(birthday) + ageYAAAAAHHHH + 1
birthday
```
    e) The number of _days_ until your next birthday.
```{r}
birthday - ymd(lubridate::today())
```
    f) The number of _months_ and _days_ until your next birthday.
```{r}
month(as.period(ymd(lubridate::today()) %--% birthday))
day(as.period(ymd(lubridate::today()) %--% birthday))
```

## Question 3
3. Suppose you have arranged for a phone call to be at 3 pm on May 8, 2015 at Arizona time. However, the recipient will be in Auckland, NZ. What time will it be there?
```{r}
with_tz(mdy_hm("May 8, 2015 3:00 pm", tz='US/Arizona'), tzone='NZ')
```

## Question 5
5. It turns out there is some interesting periodicity regarding the number of births on particular days of the year.
    a. Using the `mosaicData` package, load the data set `Births78` which records the number of children born on each day in the United States in 1978. Because this problem is intended to show how to calculate the information using the `date`, remove all the columns *except* `date` and `births`. 
```{r}
yaboiiii <- mosaicData::Births78 %>%
  select(c("date", "births"))
head(yaboiiii)
```
    b. Graph the number of `births` vs the `date` with date on the x-axis. What stands out to you? Why do you think we have this trend?
```{r}
ggplot(yaboiiii) +
  geom_point(aes(x=date, y=births))
```
It seems there are two main groups, like its bimodal. I see no reason why it would be more popular to give birth on a certain day of the week, but judging by the next question, I guess that's what we're goin with.
    c. To test your assumption, we need to figure out the what day of the week each observation is. Use `dplyr::mutate` to add a new column named `dow` that is the day of the week (Monday, Tuesday, etc). This calculation will involve some function in the `lubridate` package and the `date` column.
```{r}
yaboiiii <- yaboiiii %>%
  dplyr::mutate(dow = wday(date, label = TRUE))
head(yaboiiii)
```
    d. Plot the data with the point color being determined by the day of the week variable.
```{r}
ggplot(yaboiiii) +
  geom_point(aes(x=date, y=births, color=dow))
```